Data Processing Strategies to Determine Maximum Oxygen Uptake: A Systematic Scoping Review and Experimental Comparison with Guidelines for Reporting

### 2023-07-04 scripts/comparison.Rmd
Add plots
revision and congress version

library(purrr)
library(spiro)
library(tidyr)
library(ggplot2)
library(scales)
library(MetBrewer)

@@

## Normalize results

@@

## Turn into tidy format

@@

## Plot: Comparison of processing strategies

@@

p_comp <- ggplot(res_tidy, aes(x = parameter, y = perdiff, colour = type, fill = type)) +
  geom_hline(
    yintercept = 0,
    colour = "grey",
    alpha = 0.3
  ) +
  stat_summary(geom = "line", fun = "median") +
  stat_summary(
    geom = "ribbon", 
    fun.data = "quant", 
    fun.args = list(lower = 0.1, upper = 0.9),
    alpha = 0.1, 
    colour = NA
  ) +
  scale_x_continuous(
    name = "Parameter",
    breaks = c(15, 30, 45, 60)
  ) +
  scale_y_continuous(
    name = "% to Reference",
    labels = scales::percent,
    limits = c(-0.053,0.1)
  ) +
  scale_color_manual(
    name = "Processing Strategy",
    values = met.brewer("Java", 3)
  ) +
  scale_fill_manual(
    name = "Processing Strategy",
    values = met.brewer("Java", 3)
  ) +
  theme_classic() +
  theme(
    legend.position = c(0.8,0.8)
  )

@@

knitr::include_graphics("../plots/comparison.png")
```

## Plot: Respiratory Rate over time

```{r rr}
read_rr <- function(file) {
  dat <- spiro(file)
  # Zero-lag Butterworth filter with default parameters
  rr <- spiro_smooth(dat, smooth = "fz", columns = "RR")[, 1]
  time <- attr(dat, "raw")$time
  load <- attr(dat, "raw")$load

@@

  rr_df <- data.frame(
    id = id,
    time = time,
    value = rr,
    load = load,
    is_end = FALSE
  )

@@

# Line plots with large data are very slow in ggplot2
# using a scatter plot instead
p_rr <- ggplot(rr_all, aes(x = time, y = value, colour = is_end, alpha = is_end)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  scale_colour_manual(values = c("black", met.brewer("Peru2")[2])) +
  scale_alpha_manual(values = c(0.05, 0.3)) +
  scale_x_continuous(
    name = "Time (min)",
    limits = c(60,950),
    breaks = c(180, 360, 540, 720, 900),
    labels = c(2, 5, 8, 12, 15),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste("Respiratory Rate ", (min^-1))),
    breaks = c(20,40,60,80)
  ) +
  theme_classic()

# ggsave(filename = "../plots/rr.png", plot = p_rr, width = 6, height = 4, dpi = 600, bg = "white")

knitr::include_graphics("../plots/rr.png")
```

# iterate over all tests
fzmax <- purrr::map_dbl(seq_along(files), get_fzmax)

# get median and quantile values
median(fzmax_norm)
quant(fzmax_norm, 0.1, 0.9)
```

@@

knitr::include_graphics("../plots/example.png")
```

>>>>

library(purrr)
library(spiro)
library(tidyr)
library(ggplot2)
library(scales)
library(MetBrewer)
library(patchwork)
library(lmerTest)
library(broom.mixed)
library(multcomp)
library(ggdist)

@@

## Plot: Comparison of various strategies

### Normalize results

@@

### Turn into tidy format

@@

### Create comparison plot

@@

p_comp_pre <- ggplot(res_tidy, aes(x = parameter, y = perdiff, colour = type, fill = type, linetype = type)) +
  geom_hline(
    yintercept = 0,
    colour = "grey",
    alpha = 0.3
  ) +
  stat_summary(geom = "line", fun = "median") +
  stat_summary(
    geom = "ribbon", 
    fun.data = "quant", 
    fun.args = list(lower = 0.1, upper = 0.9),
    alpha = 0.1, 
    colour = NA
  ) +
  scale_color_manual(
    name = "Processing Strategy",
    values = met.brewer("Java", 3)
  ) +
  scale_fill_manual(
    name = "Processing Strategy",
    values = met.brewer("Java", 3)
  ) +
  scale_linetype_discrete(
    name = "Processing Strategy"
  )

p_comp <- p_comp_pre +
  scale_x_continuous(
    name = "Parameter",
    breaks = c(15, 30, 45, 60)
  ) +
    scale_y_continuous(
    name = "% to Reference",
    labels = scales::percent,
    limits = c(-0.053,0.1)
  ) +
  theme_classic() +
  theme(
    legend.position = c(0.8,0.8)
  )

@@

# save plot as EPS
# ggsave("../plots/comparison.pdf", plot = p_comp, width = 6, height = 4, dpi = 1200, bg = "white")

# create plot with different labels

p_comp_duration <- p_comp_pre +
  scale_x_continuous(
    name = "calculation interval length (seconds or breaths)",
    breaks = c(15, 30, 45, 60)
  ) +
  scale_y_continuous(
    name = expression(paste(dot(V), O[2][max], " (% to reference)")),
    labels = scales::percent,
    limits = c(-0.053,0.1)
  ) +
  theme_classic() +
  theme(
    legend.position = c(0.8,0.8)
  )
# ggsave("../plots/comparison_xlabel.png", plot = p_comp_duration, width = 6, height = 4, dpi = 600, bg = "white")
# ggsave("../plots/comparison_xlabel.pdf", plot = p_comp_duration, width = 6, height = 4, dpi = 1200, bg = "white")

knitr::include_graphics("../plots/comparison_xlabel.png")
```

## Plot: Respiratory rate and VO2 over time

```{r rr}
read_rr <- function(file) {
  dat <- spiro(file)
  # Zero-lag Butterworth filter with default parameters
  rr <- spiro_smooth(dat, smooth = "f", columns = "RR")[, 1]
  vo2 <- spiro_smooth(dat, smooth = "f", columns = "VO2")[, 1] / attr(dat, "info")$weight
  time <- attr(dat, "raw")$time
  load <- attr(dat, "raw")$load

@@

  rr_df <- data.frame(
    id = id,
    time = time,
    value_rr = rr,
    value_vo2 = vo2,
    load = load,
    is_end = FALSE
  )

@@

# Line plots with large data are very slow in ggplot2
# using a scatter plot instead
p_rr <- ggplot(rr_all, aes(x = time, y = value_rr, colour = is_end, alpha = is_end)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  scale_colour_manual(values = c("black", met.brewer("Peru2")[2])) +
  scale_alpha_manual(values = c(0.05, 0.3)) +
  scale_x_continuous(
    name = NULL,
    limits = c(60,950),
    breaks = c(180, 360, 540, 720, 900),
    labels = c(2, 5, 8, 12, 15),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste("Respiratory Rate ", (min^-1))),
    breaks = c(20,40,60,80)
  ) +
  theme_classic()

p_vo2 <- ggplot(rr_all, aes(x = time, y = value_vo2, colour = is_end, alpha = is_end, group = id)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  scale_colour_manual(values = c("black", met.brewer("Peru2")[2])) +
  scale_alpha_manual(values = c(0.05, 0.3)) +
  scale_x_continuous(
    name = "Time (min)",
    limits = c(60,950),
    breaks = c(180, 360, 540, 720, 900),
    labels = c(2, 5, 8, 12, 15),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste(dot(V), O[2], " uptake ", (ml%.%min^-1%.%kg^-1)))
  ) +
  theme_classic()

p_all <- p_rr / p_vo2 + plot_annotation(tag_levels = 'A')

# ggsave(filename = "../plots/timecourse.png", plot = p_all, width = 6, height = 7, dpi = 600, bg = "white")
# ggsave(filename = "../plots/timecourse.pdf", plot = p_all, width = 6, height = 7, dpi = 1200, bg = "white")

p_all_vert <- (p_rr | p_vo2) + plot_annotation(tag_levels = 'A')
# ggsave(filename = "../plots/timecourse_vert.png", plot = p_all_vert, width = 8, height = 3, dpi = 600, bg = "white")

knitr::include_graphics("../plots/timecourse.png")
```

@@

knitr::include_graphics("../plots/example.png")

## ECSS presentation plot
p2_raw <- ggplot(exm_tidy, aes(x = time)) +
  geom_point(
    data = exm_tidy[exm_tidy$method == "raw", ], 
    aes(y = value, alpha = method, colour = "Raw Data"),
    show.legend = FALSE
  ) +
  scale_x_continuous(
    name = "Time (min)",
    limits = c(500,820),
    breaks = c(540, 630, 720),
    labels = c(8, 10, 12),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste(dot(V), O[2], " ", (ml%.%min^-1%.%kg^-1))),
    limits = c(45,75)
  ) +
  scale_colour_manual(values = "black") +
  scale_fill_manual(
    values = "white",
    guide = "none"
  ) +  scale_alpha_manual(
    values = 0.3,
    guide = "none"
  ) +
  labs(title = "Raw data") +
  theme_classic(15) +
  theme(
    legend.position = c(0.88, 0.63),
  )

# ggsave(filename = "../plots/example2_raw.png", plot = p2_raw, width = 6, height = 4, dpi = 1200, bg = "white")

fz2 <- spiro_smooth(exm_data, "fz", "VO2")[,1]
mb2 <- spiro_smooth(exm_data, "7b", "VO2")[,1]
mt2 <- spiro_smooth(exm_data, 60, "VO2")[,1]
tb2 <- rep(max_time_binned(exm_data, 30, max = FALSE) * attr(exm_data, "info")[["weight"]], each = 30)

exm_tidy2 <- data.frame(
  method = c(rep.int("raw", length(raw_vo2)), rep.int("fz", length(fz2)), rep.int("mb", length(mb2)), rep.int("mt", length(mt2)), rep.int("tb", length(tb2))),
  value = c(raw_vo2, fz2, mb2, mt2, tb2),
  time = c(rep.int(raw_time, 3), seq_along(mt2), seq_along(tb2))
)

exm_tidy2$method <- factor(
  exm_tidy2$method, 
  levels = c("raw", "mb", "mt", "tb", "fz"), 
  labels = c("raw","Moving Breath","Moving Time","Binned Time", "Digital Filter"),
  ordered = TRUE
)
exm_tidy2$value <- exm_tidy2$value / bodymass

exm_tidy_cut <- exm_tidy2[exm_tidy2$time >= 500,]
exm_tidy_cut_copy <- exm_tidy_cut[exm_tidy_cut$method == "raw", ]
colnames(exm_tidy_cut_copy)[3] <- "timecopy"

create_gif <- function(type, title, color, savename) {
  out <- ggplot(exm_tidy_cut, aes(x = time)) +
  geom_point(
    data = exm_tidy_cut_copy, 
    aes(x = timecopy, y = value, alpha = method),
    color = "black",
    show.legend = FALSE,
    size = 1.5
  ) +
  geom_line(
    data = exm_tidy_cut[exm_tidy_cut$method %in% c(type), ], 
    aes(y = value, color = method), 
    size = 1.3,
    show.legend = FALSE,
    color = color
  ) +
  scale_x_continuous(
    name = "Time (min)",
    limits = c(500,820),
    breaks = c(540, 630, 720),
    labels = c(8, 10, 12),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste(dot(V), O[2], " ", (ml%.%min^-1%.%kg^-1))),
    limits = c(45,75)
  ) +
  scale_alpha_manual(
    values = 0.3,
    guide = "none"
  ) +
  labs(
    title = title
  ) +
  theme_classic(15) +
  theme(
    legend.position = c(0.88, 0.63),
    plot.title = element_text(colour = color)
  ) +
  gganimate::transition_reveal(time)
  gganimate::anim_save(
    paste0("../plots/", savename, ".gif"), out, fps = 25,
    res = 300, width = 6.5, height = 4.5, units = "in" # passed to grDevices::png()
  ) 
}

#create_gif("Moving Breath", "7-breath moving (Robergs & Burnett 2003)", "darkgreen", "anim1")
#create_gif("Moving Time", "60-s moving (Howley et al. 1995)", "purple", "anim2")
#create_gif("Digital Filter", "Digital Filter (Robergs et al. 2010)", "orange", "anim3")
#create_gif("Binned Time", "30-s binned", "turquoise3", "anim4")

# ECSS all methods suppl. plot

allplot <- ggplot(exm_tidy_cut, aes(x = time)) +
  geom_point(
    data = exm_tidy_cut_copy, 
    aes(x = timecopy, y = value, alpha = method),
    color = "black",
    show.legend = FALSE,
    size = 1.5
  ) +
  geom_line(
    data = exm_tidy_cut[exm_tidy_cut$method %in% c("Moving Breath", "Moving Time", "Digital Filter", "Binned Time"), ], 
    aes(y = value, color = method), 
    size = 1.3
  ) +
  scale_x_continuous(
    name = "Time (min)",
    limits = c(500,800),
    breaks = c(540, 630, 720),
    labels = c(8, 10, 12),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste(dot(V), O[2], " ", (ml%.%min^-1%.%kg^-1))),
    limits = c(45,75)
  ) +
  scale_alpha_manual(
    values = 0.3,
    guide = "none"
  ) +
  scale_color_manual(
    name = "Method",
    values = c("darkgreen", "purple", "orange", "turquoise2")
  ) +
  theme_classic(15)

#ggsave("../plots/allplot.png", plot = allplot, dpi = 300, width = 7.5, height = 4.5, bg = "white")
```

get_fmax <- function(i) {
  dta <- spiro(files[i], weight = info$bodymass[i])
  spiro_max(dta, smooth = "0.04f3")$VO2_rel
}

# iterate over all tests
fzmax <- purrr::map_dbl(seq_along(files), get_fzmax)
fmax <- purrr::map_dbl(seq_along(files), get_fmax)

# get median and quantile values
median(fzmax_norm)
#quant(fzmax_norm, 0.1, 0.9)
mean(fmax)
sd(fmax)
```

## Statistical comparison of selected strategies

```{r}
res <- read.csv("../data/results.csv")

slct <- data.frame(
  id = seq_len(nrow(res)),
  bm_07 = res$bm_7,
  bm_15 = res$bm_15,
  tb_30 = res$tb_30,
  filt = fmax,
  tm_30 = res$tm_30,
  tb_60 = res$tb_60
)

slct_long <- tidyr::pivot_longer(slct, cols = -id)

slct_long$name <- factor(
  slct_long$name, 
  levels = names(sort(apply(slct[,-1],2,mean))),
  labels = c("(f) 60-s binned", "(e) 30-s binned","(d) 30-s moving","(c) digital filter","(b) 15-breath moving","(a) 7-breath moving"))

p_selcomp <- ggplot(slct_long, aes(x = value, y = name)) +
  geom_line(aes(group = id), orientation = "y", colour = "grey80") +
  stat_dotsinterval(
    alpha = 0.7, 
    .width = c(0.5, 0.8), 
    binwidth = 0.5, dotsize = 1,
    point_interval = "mean_qi"
    ) +
  scale_x_continuous(
    name = expression(paste(dot(V), O[2][max], " ", (ml%.%min^-1%.%kg^-1))),
    expand = c(0.02,0.02),
    breaks = seq(50,75,5),
    labels = c("50","","60","","70","")
  ) +
  scale_y_discrete(
    name = NULL,
    expand = c(0.05,0.1)
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line.y = element_blank(),
    axis.text.y = element_text(colour = "black")
  )

mod <- lmer(value ~ name + (1 | id), slct_long)
anova(mod)

summary(glht(mod, linfct = mcp(name = "Tukey")), test = adjusted(type = "single-step"))

# ggsave(filename = "../plots/sel_comp.pdf", width = 8, height = 5, dpi = 600, bg = "white")
# ggsave(filename = "../plots/sel_comp.png", width = 8, height = 5, dpi = 600, bg = "white")

p_selcomp
```
